### 2023-06-29 proofs/tex/alg-pos-subtyping-lemmas.tex
wip

\begin{lemma}[Soundness of the Positive Subtyping] \label{lemma:pos-subt-soundness}
    If $[[Γ ⊢ Θ]]$, $[[Γ ⊢ iQ]]$, and $[[Γ ; Θ ⊢ uP]]$ then\\
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ us]]$
    implies $[[us : Θ|uv uP]]$ and 
    for any $[[us']]$ such that $[[Θ ⊢ us' ⇒ us]]$,
    $[[ Γ ⊢ [us']uP ≥ iQ ]]$.
\end{lemma}
\begin{proof} 
    We prove it by induction on $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ us]]$. Let us consider the last rule to infer this judgment.
    \begin{caseof}
    \item \ruleref{\ottdruleAPUVarLabel} then
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ us]]$ has shape $[[Γ;Θ ⊨ â⁺ ≥ iP' ⫤ (â⁺ :≥ iQ')]]$ where
    $[[â⁺[Δ] ∊ Θ]]$ and $[[upgrade G ⊢ iP' to Δ = iQ']]$.

    Notice that $[[â⁺[Δ] ∊ Θ]]$ and $[[Γ ⊢ Θ]]$ 
    implies $[[Γ = Δ, pnas]]$ for some $[[pnas]]$, hence, the
    soundness of upgrade (\cref{lemma:upgrade-soundness}) is applicable:
    \begin{enumerate}
        \item $[[Δ ⊢ iQ']]$ and
        \item $[[Γ ⊢ iQ ≥ iP]]$.
    \end{enumerate}

    Since $[[â⁺[Δ] ∊ Θ|uv â⁺]]$ and $[[Δ ⊢ iQ']]$, it is clear that $[[(â⁺ :≥ iQ') : Θ | uv â⁺ ]]$.

    It is left to show that $[[Γ ⊢ [us']â⁺ ≥ iP']]$ for any $[[us']]$ s.t. $[[Θ ⊢ us' ⇒ (â⁺ :≥ iQ')]]$.
    The latter weakening statement means that either $[[us']] \ni [[â⁺ :≥ iQ'']]$ or
    $[[us']] \ni [[(â⁺ :≈ iQ'')]]$ for $[[Δ ⊢ iQ'' ≥ iQ']]$. In any case,
    $[[Δ ⊢ [us']â⁺ ≥ iQ]]$. By weakening the context to $[[Γ]]$ and combining this judgment
    transitively (\cref{todo}) with $[[Γ ⊢ iQ ≥ iP]]$, we have $[[Γ ⊢ [us']â⁺ ≥ iP]]$,
    as required.

    \item \label{case:pos-subt-soundness:var} \ruleref{\ottdruleAPVarLabel}  
    then $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ us]]$ has shape $[[Γ;Θ ⊨ a⁺ ≥ a⁺ ⫤ ·]]$ .
    Then $[[uv a⁺ = ∅]]$, and $[[us]] = [[· : ·]]$ satisfies $[[us : Θ|∅]]$.
    Since $[[uv a⁺ = ∅]]$, application of any unification solution $[[us']]$ does not change $[[a⁺]]$, i.e.
    $[[ [us'] a⁺ = a⁺]]$. Therefore, $[[Γ ⊢ [us']a⁺ ≥ a⁺]]$ holds by \ruleref{\ottdruleDOneNVarLabel}.

    \item \label{case:pos-subt-soundness:shift} 
     \ruleref{\ottdruleAShiftDLabel} then
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ us]]$ has shape $[[Γ;Θ ⊨ ↓uN ≥ ↓iM ⫤ us]]$.\\
    Then the next step of the algorithm is the unification of $[[nf(uN)]]$ and $[[nf(iM)]]$.
    By the soundness of the unification algorithm (\cref{lemma:unification-soundness}),
    it returns an equivalence-only solution $[[us]]$ such that $[[us : Θ | uv uN]]$.
    By \cref{todo}, since $[[us]]$ is equivalence-only and $[[Γ ⊢ Θ]]$, $[[Θ ⊢ us' ⇒ us]]$ means 
    $[[ Γ ⊢ Sub us' ≈ Sub us : Ord uv uN ]]$ as substitutions.
    $[[ [us]nf(uN) = nf(iM) ]]$ implies $[[Γ ⊢ [us]nf(uN) ≈ nf(iM)]]$, and then 
        $[[Γ ⊢ [us']nf(uN) ≈ nf(iM)]]$. \ilyam{add lemmas}

        Let us rewrite the left-hand side and the right-hand side of $[[Γ ⊢ [us']nf(uN) ≈ nf(iM)]]$ by 
        transitivity of equivalence (\cref{corollary:equivalence-transitivity}).
        By \cref{corollary:nf-sound-wrt-subt-equiv,corollary:subst-pres-equiv},
        $[[Γ ⊢ [us']nf(uN) ≈ [us']uN ]]$. By \cref{corollary:nf-sound-wrt-subt-equiv}, 
        $[[Γ ⊢ nf(iM) ≈ iM ]]$. 
        This way, we have $[[Γ ⊢ [us']uN ≈ iM]]$.
        Then by \ruleref{\ottdruleDOneShiftULabel}
        and congruence of substitution, $[[Γ ⊢ [us']↓uN ≥ ↓iM]]$.

    \item \label{case:pos-subt-soundness:exists}
       \ruleref{\ottdruleAExistsLabel} then
        $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ us]]$ has shape $[[Γ;Θ ⊨ ∃nas.uP' ≥ ∃nbs.iQ' ⫤ us]]$ s.t. either $[[nas]]$ or $[[nbs]]$ is not empty.\\
        Then the algorithm creates fresh unification variables $[[â⁻*[Γ,nbs] ]]$, 
        substitutes the old $[[nas]]$ with them in $[[uP']]$, and makes the recursive call:
        $[[G, nbs; Θ, â⁻*[G, nbs] ⊨ [â⁻*/nas] uP ≥ iQ ⫤ us']]$, returning as the result
        $[[us]] = [[us' \ {α̂⁻*}]]$.

        Notice that $[[Γ, nbs ⊢ Θ, â⁻*[Γ, nbs] ]]$, $[[Γ,nbs ⊢ iQ']]$, and 
        $[[Γ,nbs; Θ, â⁻*[Γ, nbs] ⊢ [â⁻*/nas] uP' ]]$, so the induction hypothesis is applicable,
        that is $[[us' : Θ, â⁻*[Γ, nbs] | uv [â⁻*/nas]uP']]$ and $[[ Γ, nbs ⊢ [us'2][â⁻*/nas]uP' ≥ iQ' ]]$ for any
        $[[us'2]]$ s.t. $[[Θ, â⁻*[Γ, nbs] ⊢ us'2 ⇒ us']]$.

        Since the domain of $[[us']]$ is $[[uv [â⁻*/nas]uP']]$, the domain of 
        $[[us]] = [[us' \  {α̂⁻*}]]$ is $[[uv [â⁻*/nas]uP' \ {â⁻*}]] = [[uv ∃nas.uP']]$,
        this way, $[[us : Θ | uv ∃nas.uP']]$, as required.

        It is left to show that $[[Γ ⊢ [us2]∃nas.uP' ≥ ∃nbs.iQ']]$ for any $[[us2]]$ s.t. $[[Θ ⊢ us2 ⇒ us]]$.
        Let us consider an arbitrary such $[[us2]]$. Let us construct $[[us'2]]$, 
        extending $[[us2]]$ to the domain $[[uv [â⁻*/nas]uP']]$ with the values of $[[us']]$,
        i.e.  $[[us'2 : Θ | uv [â⁻*/nas]uP']]$, 
        and
        \[
            [[us'2(β̂±)]]  = 
            \begin{cases}
               [[us2(β̂±)]] & \text{if } [[β̂±]] \in [[uv uP']] \\
               [[us'(β̂±)]] & \text{if } [[β̂±]] \in [[â⁻*]]
            \end{cases}
        \] 
        , where the application of the unification solution to a variable returns the 
        corresponding \emph{unification entry}. 
        Notice that $[[Sub us'2|uv uP']] = [[us2]]$.
    It is easy to see that $[[Θ ⊢ us'2 ⇒ us']]$: 
    \begin{enumerate}
        \item if $[[β̂±]] \in [[uv uP']]$ then $[[us'2(β̂±)]] = [[us2(β̂±)]] \Rightarrow [[us(β̂±)]] = [[us'(β̂±)]]$;
        \item it $[[β̂±]] \in [[â⁻*]]$ then $[[us'2(β̂±)]] = [[us'(β̂±)]] \Rightarrow [[us'(β̂±)]]$,
    \end{enumerate}
    which means that the induction hypothesis can be applied to $[[us'2]]$, i.e.
    $[[ Γ, nbs ⊢ [us'2][â⁻*/nas]uP' ≥ iQ' ]]$.\\
    Notice that
    $
    \begin{aligned}[t]
                 [[ [us'2][â⁻*/nas]uP' ]] &= [[ [Sub us'2|{â⁻*} ○ â⁻*/nas][Sub us'2|uv uP']uP' ]]
                                          && \text{by substitution properties \ilyam{todo}}\\
                                          &= [[ [Sub us'2|{â⁻*} ○ â⁻*/nas][us2]uP' ]]
                                          && \text{since $[[Sub us'2|uv uP']] = [[us2]]$}.
    \end{aligned}
    $\\
    Also notice that the domain of $[[Sub us'2|{â⁻*} ○ â⁻*/nas]]$ is $[[nas]]$,
    and the range is $[[Γ, nbs]]$, i.e. $[[Γ, nbs ⊢ Sub us'2|{â⁻*} ○ â⁻*/nas : nas]]$, 
    which means that we can apply \ruleref{\ottdruleDOneForallLabel} to 
    $[[ Γ, nbs ⊢ [Sub us'2|{â⁻*} ○ â⁻*/nas][us2]uN' ≤ iM' ]]$
    to obtain $[[ Γ ⊢ [us2]∃nas.uP' ≥ ∃nbs.iQ' ]]$, as required.
    \end{caseof}
\end{proof}

\begin{lemma}[Completeness of the Positive Subtyping] \label{lemma:pos-subt-completeness}
    Suppose that $[[Γ ⊢ Θ]]$, $[[Γ ⊢ iQ]]$ and $[[Γ ; Θ ⊢ uP]]$ and
    there exists $[[us : Θ | uv uP]]$ such that $[[ Γ ⊢ [us]uP ≥ iQ ]]$.
    Then there exists $[[us']]$ such that $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ us']]$
    and $[[Θ ⊢ us ⇒ us']]$.
\end{lemma}
\begin{proof}
    We prove it by induction on $[[ Γ ⊢ [us]uP ≥ iQ ]]$.
    Let us consider the last rule used in the derivation of  $[[ Γ ⊢ [us]uP ≥ iQ ]]$,
    but first consider the base case for the substitution $[[ [us]uP ]]$:

    \begin{caseof}
        \item \label{case:pos-subt-complete-base} $[[uP]] = [[ ∃nbs.α̂⁺ ]]$ (for potentially empty $[[pbs]]$)\\
        Then by assumption, there exists $[[us : Θ | uv uP]]$ such that $[[ Γ ⊢ ∃nbs.[us]α̂⁺ ≥ iQ ]]$.
        $[[us : Θ | uv α̂⁺]]$ means that  $[[us]]$ is either $[[ (α̂⁺ :≈ iP') ]]$ or $[[ (α̂⁺ :≥ iP') ]]$,
        where $[[Δ ⊢ iP']]$ and $[[α̂⁺[Δ] ∊ Θ]]$.
        $[[ Γ ⊢ ∃nbs.[us]α̂⁺ ≥ iQ ]]$ means that $[[Γ ⊢ iP' ≥ iQ]]$
        because multiple inversions of \ruleref{\ottdruleDOneExistsLabel} 
        gives us $[[Γ ⊢ [us]α̂⁺ ≥ iQ]]$ since $[[ {nbs} ∩ fv [us]α̂⁺ = ∅]]$.

        In the algorithm, after multiple applications of \ruleref{\ottdruleAExistsLabel}
        the type $[[∃nbs.α̂⁺]]$ is reduced to $[[α̂⁺]]$.
        Next, the algorithm tries to apply
        \ruleref{\ottdruleAPUVarLabel}
        and the resulting solution is $[[us']] = [[(α̂⁺ :≥ iQ')]]$ where
        $[[upgrade Γ ⊢ iQ to Δ = iQ']]$.

        Why does the upgrade procedure terminates?
        Because $[[iP']]$ satisfies the pre-conditions of the completeness of the upgrade
        (\cref{lemma:upgrade-completeness})
        :
        \begin{itemize}
            \item $[[Δ ⊢ iP']]$ because $[[iP' = [us]α̂⁺]]$, $[[us : Θ | uv uP]]$ and 
            $[[α̂⁺[Δ] ∊ Θ | uv uP]]$,
            \item $[[Γ ⊢ iP' ≥ iQ]]$ as noted before
        \end{itemize}
        Moreover, completeness of the upgrade also gives us $[[Γ ⊢ iP' ≥ iQ']]$
        and further, we strengthen it to $[[Δ ⊢ iP' ≥ iQ']]$
        (since by the soundness of the upgrade (\cref{lemma:upgrade-soundness}),
        $[[Δ ⊢ iQ']]$).

        It means that $[[Θ ⊢ (α̂⁺ :≈ iP') ⇒ (α̂⁺ :≥ iQ') ]]$ and 
        $[[ Θ ⊢ (α̂⁺ :≥ iP') ⇒ (α̂⁺ :≥ iQ') ]]$, which means that in any case, 
        $[[ Θ ⊢ us ⇒ us']]$.

        \item \label{case:pos-subt-complete-pvar}
        $[[ Γ ⊢ [us]uP ≥ iQ ]]$ is derived by \ruleref{\ottdruleDOnePVarLabel}, 
        i.e. $[[iP]] = [[ [us]uP ]] = [[ α⁺ ]] = [[iQ]]$.
        Here the first equality holds because $[[uP]]$ is not a unification variable:
        this case has been covered by \cref{case:pos-subt-complete-base}.
        The second equality hold because \ruleref{\ottdruleDOnePVarLabel} was applied.

        Notice that $[[us : Θ | uv α⁺]]$ = $[[us : Θ | ∅]]$ = $[[·]]$.

        The algorithm applies \ruleref{\ottdruleAPVarLabel} and 
        infers $[[us']] = [[·]]$, i.e. $[[Γ;Θ ⊨ a⁺ ≥ a⁺ ⫤ ·]]$.

        Since $[[Θ ⊢ · ⇒ ·]]$, we have $[[Θ ⊢ us ⇒ us']]$.

        \item \label{case:pos-subt-complete-upshift} $[[ Γ ⊢ [us]uP ≥ iQ ]]$ is derived by \ruleref{\ottdruleDOneShiftDLabel},

        Then $[[ uP ]] = [[ ↓uN ]]$, since the substitution $[[ [us]uP ]]$ must preserve the 
        top-level constructor of $[[uP]]\neq [[α̂⁺]]$ (the case $[[uP]] = [[α̂⁺]]$ has been covered
        by \cref{case:pos-subt-complete-base}), and $[[uQ]] = [[ ↓iM ]]$,
        and by inversion, $[[ Γ ⊢ [us]uN ≈ iM ]]$.

        Since both types start with $[[↓]]$, 
        the algorithm tries to apply \ruleref{\ottdruleAShiftDLabel}: 
        $[[G;Θ ⊨ ↓uN ≥ ↓iM ⫤ us']]$. The premise of this rule is the
        unification of $[[nf(uN)]]$ and $[[nf(iM)]]$:
        $[[Γ;Θ ⊨ nf(uN) ≈u nf(iM) ⫤ us']]$. Let us show that the unification successfully 
        terminates and returns $[[us']] = [[nf(us)]]$.

        To demonstrate that the unification terminates, we apply the completeness 
        of the unification algorithm (\cref{lemma:unification-completeness}). 
        In order to do that, we need to provide a unifier of 
        $[[nf(uN)]]$ and $[[nf(iM)]]$. Thankfully, $[[nf(us)]]$ does it.

        \begin{itemize}
            \item $[[nf(uN)]]$ and $[[nf(iM)]]$ are normalized 
            \item $[[Γ ; Θ ⊢ nf(uN)]]$ because $[[Γ ; Θ ⊢ uN]]$ (\cref{todo})
            \item $[[Γ ⊢ nf(iM)]]$ because $[[Γ ⊢ iM]]$ (\cref{corollary:wf-nf})
            \item $[[nf(us) : Θ | uv nf(uN)]]$ because $[[us: Θ | uv uN]]$ (\cref{todo})
            \item $ \begin{aligned}[t]
                    [[ Γ ⊢ [us]uN ≈ iM ]] &\Rightarrow [[ [us]uN ≈ iM ]]
                                          && \text {by \cref{lemma:equiv-completeness}}\\
                                          &\Rightarrow [[ nf([us]uN) = nf(iM) ]]
                                          && \text {by \cref{lemma:normalization-completeness}}\\
                                          &\Rightarrow [[ [nf(us)]nf(uN) = nf(iM) ]]
                                          && \text {by \cref{lemma:norm-subst-distr}}\\
                    \end{aligned}
                  $
        \end{itemize}
        Then by the completeness of the unification,
        $[[Γ ; Θ ⊨ nf(uN) ≈u nf(iM) ⫤ nf(us)]]$.
        This way, the subtyping algorithm terminates and the resulting solution is
        $[[nf(us)]]$.

        It is left to note that $[[Θ ⊢ us ⇒ nf(us)]]$, by \cref{todo}, since $[[Θ ⊢ us ≈ nf(us)]]$

>>>>

\begin{lemma}[Soundness of the Positive Subtyping] \label{lemma:pos-subt-soundness}
    If $[[Γ ⊢ Θ]]$, $[[Γ ⊢ iQ]]$, $[[Γ ; Θ ⊢ uP]]$, and 
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$,
    then $[[Θ ⊢ SC]]$ and
    for any normalized $[[uσ]]$ such that $[[Θ ⊢ uσ : SC]]$,
    $[[ Γ ⊢ [uσ]uP ≥ iQ ]]$.
\end{lemma}
\begin{proof} 
    We prove it by induction on $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$. 
    Let us consider the last rule to infer this judgment.
    \begin{caseof}
    \item \ruleref{\ottdruleAPUVarLabel} then
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ has shape $[[Γ;Θ ⊨ â⁺ ≥ iP' ⫤ (â⁺ :≥ iQ')]]$ where
    $[[â⁺[Δ] ∊ Θ]]$ and $[[upgrade G ⊢ iP' to Δ = iQ']]$.

    Notice that $[[â⁺[Δ] ∊ Θ]]$ and $[[Γ ⊢ Θ]]$ 
    implies $[[Γ = Δ, pnas]]$ for some $[[pnas]]$, hence, the
    soundness of upgrade (\cref{lemma:upgrade-soundness}) is applicable:
    \begin{enumerate}
        \item $[[Δ ⊢ iQ']]$ and
        \item $[[Γ ⊢ iQ' ≥ iP]]$.
    \end{enumerate}

    Since $[[â⁺[Δ] ∊ Θ]]$ and $[[Δ ⊢ iQ']]$, 
    it is clear that $[[ Θ ⊢ (â⁺ :≥ iQ') ]]$.

    It is left to show that $[[Γ ⊢ [uσ]â⁺ ≥ iP']]$ for any normalized $[[uσ]]$ 
    s.t. $[[Θ ⊢ uσ : (â⁺ :≥ iQ')]]$.
    The latter means that $[[ Θ(â⁺) ⊢ [uσ]â⁺ ≥ iQ' ]]$, i.e. $[[Δ ⊢ [uσ]â⁺ ≥ iQ']]$. 
    By weakening the context to $[[Γ]]$ and combining this judgment
    transitively with $[[Γ ⊢ iQ' ≥ iP]]$, we have $[[Γ ⊢ [uσ]â⁺ ≥ iP]]$,
    as required.

    \item \label{case:pos-subt-soundness:var} \ruleref{\ottdruleAPVarLabel}  
    then $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ has shape $[[Γ;Θ ⊨ a⁺ ≥ a⁺ ⫤ ·]]$ .
    Then $[[uv a⁺ = ∅]]$, and $[[SC]] = [[·]]$ satisfies $[[Θ ⊢ ·]]$.
    Since $[[uv a⁺ = ∅]]$, application of any substitution $[[uσ]]$ 
     does not change $[[a⁺]]$, i.e. $[[ [uσ] a⁺ = a⁺]]$.
      Therefore, $[[Γ ⊢ [uσ]a⁺ ≥ a⁺]]$ holds by \ruleref{\ottdruleDOneNVarLabel}.

    \item \label{case:pos-subt-soundness:shift} 
     \ruleref{\ottdruleAShiftDLabel} then
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ has shape $[[Γ;Θ ⊨ ↓uN ≥ ↓iM ⫤ SC]]$.\\
    Then the next step of the algorithm is the unification of $[[nf(uN)]]$ and $[[nf(iM)]]$,
    and it returns the resulting unification constraint $[[lift UC]] = [[SC]]$ as the result.
    By the soundness of unification (\cref{lemma:unification-soundness}),
    $[[Θ ⊢ SC]]$ and for any normalized $[[uσ]]$, $[[Θ ⊢ uσ : SC]]$
    implies $[[ [uσ]nf(uN) = nf(iM) ]]$, 
    then we rewrite the left-hand side by \cref{lemma:norm-subst-distr}:
    $[[ nf([uσ]uN) = nf(iM) ]]$ and apply \cref{lemma:subt-equiv-algorithmization}:
    $[[Γ ⊢ [uσ]uN ≈ iM]]$, then by \ruleref{\ottdruleDOneShiftULabel},
    $[[Γ ⊢ ↓[uσ]uN ≥ ↓iM]]$.

    \item \label{case:pos-subt-soundness:exists}
       \ruleref{\ottdruleAExistsLabel} then
        $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ has shape $[[Γ;Θ ⊨ ∃nas.uP' ≥ ∃nbs.iQ' ⫤ SC]]$ s.t. either 
        $[[nas]]$ or $[[nbs]]$ is not empty.\\
        Then the algorithm creates fresh unification variables $[[â⁻*[Γ,nbs] ]]$, 
        substitutes the old $[[nas]]$ with them in $[[uP']]$, and makes the recursive call:
        $[[G, nbs; Θ, â⁻*[G, nbs] ⊨ [â⁻*/nas] uP' ≥ iQ' ⫤ SC']]$, returning as the result
        $[[SC]] = [[SC' \ {α̂⁻*}]]$.

        Let us take an arbitrary normalized $[[uσ]]$ s.t. $[[Θ ⊢ uσ : SC' \ {α̂⁻*}]]$.
        We wish to show $[[Γ ⊢ [uσ]uP ≥ iQ]]$, i.e. $[[Γ ⊢ ∃nas.[uσ]uP' ≥ ∃nbs.iQ']]$.
        To do that, we apply \ruleref{\ottdruleDOneExistsLabel}, and what is left to show is
        $[[Γ, nbs ⊢ [iNs/nas][uσ]uP' ≥ iQ']]$ for some $[[iNs]]$.
        If we construct a normalized $[[uσ']]$ such that $[[Θ, â⁻*[G, nbs] ⊢ uσ' : SC']]$
        and for some $[[iNs]]$, $[[ [iNs/nas][uσ]uP' ]] = [[ [uσ'][â⁻*/nas]uP' ]]$,
        we can apply the induction hypothesis to 
        $[[Γ, nbs; Θ, â⁻*[G, nbs] ⊨ [â⁻*/nas] uP ≥ iQ ⫤ SC']]$ and infer 
        the required subtyping.

        Let us construct such $[[uσ']]$ by extending $[[uσ]]$ with $[[â⁻*]]$
        mapped to the corresponding types in $[[SC']]$:
        $$
        [[ [uσ']β̂± ]]  = 
            \begin{cases}
               [[ [uσ]β̂± ]] & \text{if } [[β̂±]] \in [[dom(SC') \ {α̂⁻*}]]  \\
               [[ nf(iN) ]] & \text{if } [[β̂±]] \in [[α̂⁻*]] \text{ and } [[(β̂± :≈ iN)]] \in SC' \\
            \end{cases}
        $$

        It is easy to see that $[[uσ']]$ is normalized.
        Let us show that $[[Θ, â⁻*[G, nbs] ⊢ uσ' : SC']]$.
        Let us take an arbitrary entry $[[scE]]$ from $[[SC']]$ restricting a variable $[[β̂±]]$.
        Suppose $[[β̂±]] \in [[dom(SC') \ {α̂⁻*}]]$. Then
        $[[ (Θ, â⁻*[G, nbs])(β̂±) ⊢ [uσ']β̂± : scE ]]$ is
        rewritten as $[[ Θ(β̂±) ⊢ [uσ]β̂± : scE ]]$, which holds since $[[Θ ⊢ uσ : SC']]$.
        Suppose $[[β̂±]] = [[αî⁻]] \in [[α̂⁻*]]$. Then
        $[[scE]] = [[(αî⁻ :≈ iN)]]$ for some $[[iN]]$, 
        $[[ [uσ']αî⁻ ]] = [[ nf(iN) ]]$ by the definition,
        and $[[ Γ, nbs ⊢ nf(iN) : (αî⁻ :≈ iN) ]]$ by \ruleref{\ottdruleSATSCENEqLabel},
        since $[[Γ ⊢ nf(iN) ≈ iN]]$ by \cref{lemma:subt-equiv-algorithmization}.

        Finally, let us show that $[[ [iNs/nas][uσ]uP' ]] = [[ [uσ'][â⁻*/nas]uP' ]]$.
        For $[[iNi]]$, we take the \emph{normalized} type restricting $[[αî⁻]]$ in $[[SC']]$.
        Let us take an arbitrary variable from $[[uP]]$.
        \begin{enumerate}
            \item If this variable is a unification variable $[[β̂±]]$, then
                $[[ [iNs/nas][uσ] β̂± ]] = [[ [uσ]β̂± ]] $, since $[[Θ ⊢ uσ : SC' \ {α̂⁻*}]]$ and 
                $[[ dom(Θ) ∩ {nas} = ∅ ]]$.

                Notice that $[[β̂±]] \in [[dom(Θ)]]$, which is disjoint from $[[{α̂⁻*}]]$, 
                that is $[[β̂±]] \in [[dom(SC') \ {α̂⁻*}]]$. This way,
                $[[ [uσ'][â⁻*/nas]β̂± ]] = [[  [uσ']β̂± ]] = [[ [uσ]β̂± ]]$ by the definition 
                of $[[uσ']]$,
            \item If this variable is a regular variable $[[β±]] \notin [[nas]]$, then 
                $[[ [iNs/nas][uσ] β± ]] = [[ β± ]] $ and $[[ [uσ'][â⁻*/nas]β± ]] = [[ β± ]]$. 
            \item If this variable is a regular variable $[[αi⁻]] \in [[nas]]$, then 
                $[[ [iNs/nas][uσ] αi⁻ ]] = [[ iNi ]] = [[ nf(iNi) ]]$
                (the latter equality holds since $[[iNi]]$ is normalized)
                and $[[ [uσ'][â⁻*/nas]αi⁻ ]] = [[  [uσ']αî⁻ ]] = [[ nf(iNi) ]]$.
        \end{enumerate}
    \end{caseof}
\end{proof}

\begin{lemma}[Completeness of the Positive Subtyping] \label{lemma:pos-subt-completeness}
    Suppose that $[[Γ ⊢ Θ]]$, $[[Γ ⊢ iQ]]$ and $[[Γ ; Θ ⊢ uP]]$.
    Then for any $[[Θ ⊢ uσ]]$ such that $[[ Γ ⊢ [uσ]uP ≥ iQ ]]$,
    there exists $[[Γ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ such that $[[Θ ⊢ SC]]$ and moreover, $[[Θ ⊢ uσ : SC]]$.
\end{lemma}
\begin{proof}
    Let us prove this lemma by induction on $[[ Γ ⊢ [uσ]uP ≥ iQ ]]$.
    Let us consider the last rule used in the derivation,
    but first, consider the base case for the substitution $[[ [uσ]uP ]]$:
    \begin{caseof}
        \item \label{case:pos-subt-complete-base} $[[uP]] = [[ ∃nbs.α̂⁺ ]]$ (for potentially empty $[[nbs]]$)\\
        Then by assumption, $[[ Γ ⊢ ∃nbs.[uσ]α̂⁺ ≥ iQ ]]$ (where $[[ {nbs} ∩ fv [uσ]α̂⁺ = ∅]]$).
        By \cref{todo}, $[[ Γ ⊢ ∃nbs.[uσ]α̂⁺ ≥ iQ ]]$ means 
        $[[ Γ ⊢ [iNs/nbs][uσ]α̂⁺ ≥ iQ ]]$, and hence $[[ Γ ⊢ [uσ]α̂⁺ ≥ iQ ]]$.
        By inversion, $[[Γ ; Θ ⊢ ∃nbs.α̂⁺]]$ implies $[[â⁺[Δ] ∊ Θ]]$ for some $[[Δ]]$.

        In the algorithm trying to infer the subtyping 
        $[[Γ; Θ ⊨ ∃nbs.α̂⁺ ≥ iQ ⫤ SC]]$,
        after multiple applications of \ruleref{\ottdruleAExistsLabel}
        the type $[[∃nbs.α̂⁺]]$ is reduced to $[[α̂⁺]]$.
        Next, the algorithm tries to apply
        \ruleref{\ottdruleAPUVarLabel}
        and the resulting restriction is $[[SC']] = [[(α̂⁺ :≥ iQ')]]$ where
        $[[upgrade Γ ⊢ iQ to Δ = iQ']]$.

        Why does the upgrade procedure terminates?
        Because $[[ [uσ]α̂⁺ ]]$ satisfies the pre-conditions of the completeness of the upgrade
        (\cref{lemma:upgrade-completeness}):
        \begin{itemize}
            \item $[[Δ ⊢ iP']]$ because $[[iP' = [uσ]α̂⁺]]$ and $[[Θ ⊢ uσ]]$ and 
            $[[α̂⁺[Δ] ∊ Θ]]$,
            \item $[[Γ ⊢ iP' ≥ iQ]]$ as noted above
        \end{itemize}
        Moreover, the completeness of the upgrade also gives us $[[Γ ⊢ iP' ≥ iQ']]$
        and further, we strengthen it to $[[Δ ⊢ iP' ≥ iQ']]$
        (since by the soundness of the upgrade (\cref{lemma:upgrade-soundness}),
        $[[Δ ⊢ iQ']]$).  It means that $[[Δ ⊢ iP' : (α̂⁺ :≥ iQ') ]]$, that is 
        $[[Θ ⊢ uσ : (α̂⁺ :≥ iQ') ]]$, as required.

        \item \label{case:pos-subt-complete-pvar}
        $[[ Γ ⊢ [uσ]uP ≥ iQ ]]$ is derived by \ruleref{\ottdruleDOnePVarLabel}, 
        i.e. $[[iP]] = [[ [uσ]uP ]] = [[ α⁺ ]] = [[iQ]]$.
        Here the first equality holds because $[[uP]]$ is not a unification variable:
        it has been covered by \cref{case:pos-subt-complete-base}.
        The second equality hold because \ruleref{\ottdruleDOnePVarLabel} was applied.

        The algorithm applies \ruleref{\ottdruleAPVarLabel} and 
        infers $[[SC]] = [[·]]$, i.e. $[[Γ;Θ ⊨ a⁺ ≥ a⁺ ⫤ ·]]$.
        Then $[[Θ ⊢ uσ : ·]]$ holds trivially.

        \item \label{case:pos-subt-complete-upshift} $[[ Γ ⊢ [uσ]uP ≥ iQ ]]$ 
        is derived by \ruleref{\ottdruleDOneShiftDLabel},

        Then $[[ uP ]] = [[ ↓uN ]]$, since the substitution $[[ [uσ]uP ]]$ must preserve the 
        top-level constructor of $[[uP]]\neq [[α̂⁺]]$ (the case $[[uP]] = [[α̂⁺]]$ has been covered
        by \cref{case:pos-subt-complete-base}), and $[[uQ]] = [[ ↓iM ]]$,
        and by inversion, $[[ Γ ⊢ [uσ]uN ≈ iM ]]$.

        Since both types start with $[[↓]]$, 
        the algorithm tries to apply \ruleref{\ottdruleAShiftDLabel}: 
        $[[G;Θ ⊨ ↓uN ≥ ↓iM ⫤ SC]]$. The premise of this rule is the
        unification of $[[nf(uN)]]$ and $[[nf(iM)]]$:
        $[[Γ;Θ ⊨ nf(uN) ≈u nf(iM) ⫤ UC]]$. And the algorithm 
        returns it as a subtyping constraint $[[SC]] = [[lift UC]]$.

        To demonstrate that the unification terminates
        ant $[[uσ]]$ satisfies the resulting constraints, 
        we apply the completeness 
        of the unification algorithm (\cref{lemma:unification-completeness}). 
        In order to do that, we need to provide a substitution unifying  
        $[[nf(uN)]]$ and $[[nf(iM)]]$. 
        Let us show that $[[nf(uσ)]]$ is such a substitution.

        \begin{itemize}
            \item $[[nf(uN)]]$ and $[[nf(iM)]]$ are normalized 
            \item $[[Γ ; Θ ⊢ nf(uN)]]$ because $[[Γ ; Θ ⊢ uN]]$ (\cref{todo})
            \item $[[Γ ⊢ nf(iM)]]$ because $[[Γ ⊢ iM]]$ (\cref{corollary:wf-nf})
            \item $[[ Θ ⊢ nf(uσ) ]]$ because $[[Θ ⊢ uσ ]]$ (\cref{todo})
            \item $ \begin{aligned}[t]
                    [[ Γ ⊢ [uσ]uN ≈ iM ]] &\Rightarrow [[ [uσ]uN ≈ iM ]]
                                          && \text {by \cref{lemma:equiv-completeness}}\\
                                          &\Rightarrow [[ nf([uσ]uN) = nf(iM) ]]
                                          && \text {by \cref{lemma:normalization-completeness}}\\
                                          &\Rightarrow [[ [nf(uσ)]nf(uN) = nf(iM) ]]
                                          && \text {by \cref{lemma:norm-subst-distr}}\\
                    \end{aligned}
                  $
        \end{itemize}
        Then by the completeness of the unification,
        $[[Γ ; Θ ⊨ uN ≈u iM ⫤ UC]]$ exists, and
        $[[Θ ⊢ nf(uσ) : lift UC]]$.  Then by \cref{todo}, $[[Θ ⊢ uσ : lift UC]]$.

      \item $[[ Γ ⊢ [uσ]uP ≥ iQ ]]$ is derived by \ruleref{\ottdruleDOneExistsLabel}.\\
      We should only consider the case
      when the substitution $[[ [uσ]uP ]]$ results in the existential type 
      $[[∃nas.iP'']]$ (for $[[iP'']] \neq [[∃]]\dots$) by congruence, 
      i.e. $[[uP = ∃nas.uP']]$ (for $[[uP']] \neq [[∃]]\dots$) and $[[ [us]uP' = iP'' ]]$.
      This is because the case when $[[uP = ∃nbs.α̂⁺]]$ has been covered
      (\cref{case:pos-subt-complete-base}), and thus, the substitution $[[uσ]]$ must
      preserve all the outer quantifiers of $[[uP]]$ and does not generate any new ones.
